---
title: "PCoA"
author: "Mia Akbar"
date: "`r Sys.Date()`"
output: pdf_document
---

The following analysis (and related notes) were informed by Austen et al., (2014) and their detailed code.

```{r setup}
library(tidyr)
library(dplyr)
library(ggplot2)

theme_custom<-function(base_size = 24, base_family = ""){
  theme_bw(base_size = base_size, base_family = base_family) %+replace% 
    theme(axis.text = element_text(size = rel(0.8),colour="black"),
          axis.ticks = element_line(colour = "black"), 
          legend.position = "none",
          panel.border = element_rect(fill = NA, colour = NA), 
          panel.grid.major = element_line(colour = NA,size = 0), 
          panel.grid.minor = element_line(colour = NA,size = 0), 
          axis.line = element_line(colour ="black")
    )
}

setwd("./")
```

# Prepare data

Relative floral display data, seed family/population information and additional traits 

```{r raw_data}
Data <- read.csv("./Data/LythrumDisplay2008.csv")
names(Data)[grep("Num",names(Data))]<-"Ind"
str(Data)

# Veg: Final vegetative height
# PC1: Position on axis characterizing FTS trade-off
# HinfW: Inflorescence biomass at harvest
```

Remove problematic individuals 

```{r cancelled_guys}
Data <- Data %>% 
  filter(TotFlwr > 10) %>% 
  # Display too small to meaningfully characterize schedule shape
  filter(Ind != 187 & Ind != 440) 
# Start day not recorded 
```

Long format of floral display data

```{r display_long}
Disp <- Data %>% 
  pivot_longer(d0:d60, 
               names_to = "RelDay", 
               values_to = "OpFlwr") %>% 
  mutate(RelDay = as.numeric(gsub("d","",RelDay)),
         AbsDay = RelDay + Start) %>% 
  select(Ind, Fam, Pop, Lat, AbsDay, OpFlwr)
```

Plant family and population information

```{r pop_info}
PopInfo <- Data %>% 
  select(Ind, Fam, Pop, Lat)
PopInfo <- PopInfo %>% mutate(RowNum= row_number()) # for combining w PCoA axes later
```

Convert Disp to wide form in order to make a K-S matrix 

```{r wide_disp}
WideDat <- pivot_wider(Disp, id_cols= Ind, names_from = AbsDay, values_from = OpFlwr)
WideDat <- WideDat[, order(as.numeric(colnames(WideDat)))]
Dat <- left_join(WideDat, PopInfo, by="Ind")
Dat <- Dat %>% dplyr::select(Ind, Fam, Pop, Lat, everything())
Dat[is.na(Dat)] <- 0
```

```{r matrix_prep}
no_plants<-369	#number of plants (rows)
first_flcol<-5	#first column containing flower count observations
last_flcol<-119	#last column containing flower count observations
first_flday<-8	#first day of flowering
#If column of plant ids is not already behaving as a character, tell it to do so
Dat[,1]<-as.character(Dat[,1])
```

# Part II: KS Distances 

Create no_plants vectors (one per plant), each of length = total number of flowers produced by that plant. 

```{r generating_dayvector}
#We will first require a vector reporting the days on which flowers were counted. 
days<-as.numeric(names(Dat))[first_flcol:last_flcol]
#We need a parallel vector reporting the columns that correspond to each day
columns<-c(first_flcol:last_flcol)

#Run loops to create the vectors, saving each as an element in a list (named "flowerlist")  
flowerlist<-vector(mode="list", length=no_plants)

for (n in 1:no_plants) {
for (p in 1:length(days))	 {

vect<-if (days[p]==first_flday){
	rep(x=days[p], times=Dat[n,columns[p]])} else {
	c(vect, rep(x=days[p], times=Dat[n,columns[p]]))}
	
	}	#end loop over days of flower counts
	
	flowerlist[[n]]<-c(vect)
	}	#end loop over plants
```


```{r computing_ksdistance}
#Kolmogorov-Smirnov distance matrix
#The pairwise k-s distance measured in R is sensitive to differences in locations. Therefore, to calculate k-s distances, we want to standardize each individual's flowering duration to run from 0 (first day flowering) to 1 (last day flowering).  This removes differences in duration, onset, and end from our comparison of schedule shape. 

list.std<-vector(mode="list", length=no_plants)
for (n in 1:no_plants) {
	list.std[[n]]<-((flowerlist[[n]]-min(flowerlist[[n]]))/(max(flowerlist[[n]])-min(flowerlist[[n]])))
	}
```

```{r build_matrix}
#Build the K-S distance matrix, indexing through each pair ij of plants.
ks<-matrix(nrow=no_plants, ncol=no_plants)

for (i in 1:no_plants) {

for (j in 1:no_plants)	{
	
	ks.ij<-ks.test(list.std[[i]], list.std[[j]])
	ks[i,j]<-ks.ij$statistic
	
	}}
```

# PART III: PCoA (constrainted ordination)   

```{r load_ks, warning=FALSE, message=FALSE}
ks <- read.csv("./Data/KSMatrix.csv") #only necessary if you did not just compute the matrix 
ks_pcoa<-cmdscale(d=ks, k=10, eig=TRUE, add=FALSE, x.ret=TRUE)	
```

```{r pcoa_eig}
ks_pcoa$eig
ks_pcoa$eig/sum(ks_pcoa$eig)
```

Here, the first four axes capture 41%, 20%, 12% and 9.8% of variation, respectively. NOTE- however, that these values are probably inflated by negative eigenvalues of later axes

K-S distances are non-metric.  How much variation ends up on negative axes (i.e., how much variation cannot be captured in Euclidean space)

```{r pcoa_var_all}
sum(ks_pcoa$eig[which(ks_pcoa$eig<0)])/sum(ks_pcoa$eig[which(ks_pcoa$eig>0)])
```

The total magnitude of variation on negative axes is 43% of that on positive axes. This is certainly inflating our above estimate of variation on the first few axes.

We can adjust our estimates of proportion variation explained by first axes by including only positive eigenvalues in the denominator

```{r pcoa_var_pos}
ks_pcoa$eig[1:5]/sum(ks_pcoa$eig[which(ks_pcoa$eig>0)])
```

This changes our estimate of variation on the first few axes to 23.5%, 11.5%, 6.9%, etc.

# Plotting 

ifelse(!dir.exists(file.path("./", "Figures")), dir.create(file.path("./", "Figures")), FALSE)

## Bivariate PCoA1 & PCoA2

```{r join_popinfo}
PlotPcoa <- as.data.frame(ks_pcoa$points)
PlotPcoa <- PlotPcoa %>% mutate(RowNum = row_number())
PlotPcoa <- left_join(PlotPcoa, PopInfo, by = "RowNum")
```

```{r scatterplot}
pcoa <- ggplot(data = PlotPcoa, aes(x=V1, y=V2, colour=Lat)) + 
  geom_point(size=3, alpha=0.7) + 
  theme_custom() + 
  xlab("PCoA1 (23.5%)") + 
  ylab("PCoA2(11.6%)") + 
  scale_color_viridis(option="magma",direction = -1) + 
  ylim(-0.45,0.45) + 
  xlim(-0.45,0.45)
```

```{r export}
pdf("./Figures/PCoA/PCoA1xPCoA2.pdf",width=10,height=6)
pcoa
dev.off()
```

## Histograms of representative individuals

### Bottom left

```{r hist_bl}
BL <- Disp %>% filter(Num=="27") %>% 
  ggplot(aes(x=AbsDay, y=OpFlwr/TotFlwr)) + geom_col() + 
  theme_classic() + 
  theme(axis.text.x=element_blank(), 
      axis.ticks.x=element_blank(), 
      axis.text.y=element_blank(), 
      axis.ticks.y=element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()) 
pdf("./Figures/PCoA/Ind27Hist.pdf",width=10,height=6)
BL
dev.off()
```


### Top left

```{r hist_tl}
TL <- Flwr %>% filter(Num=="285") %>% 
  ggplot(aes(x=AbsDay, y=OpFlwr/TotFlwr)) + geom_col() + 
  theme_classic() + 
  theme(axis.text.x=element_blank(), 
      axis.ticks.x=element_blank(), 
      axis.text.y=element_blank(), 
      axis.ticks.y=element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()) 
pdf("./Figures/PCoA/Ind285Hist.pdf",width=10,height=6)
TL
dev.off()
```

### Top right 

```{r hist_tr}
TR <- Flwr %>% filter(Num=="83") %>% 
  ggplot(aes(x=AbsDay, y=OpFlwr/TotFlwr)) + geom_col() + 
  theme_classic() + 
  theme(axis.text.x=element_blank(), 
      axis.ticks.x=element_blank(), 
      axis.text.y=element_blank(), 
      axis.ticks.y=element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()) 
pdf("./Figures/PCoA/Ind83Hist.pdf",width=10,height=6)
TR
dev.off()
```

### Bottom right 

```{r hist_br}
BR <- Flwr %>% filter(Num=="319") %>% ggplot(aes(x=AbsDay, y=OpFlwr/TotFlwr)) + geom_col() + theme_classic() + theme(axis.text.x=element_blank(), 
      axis.ticks.x=element_blank(), 
      axis.text.y=element_blank(), 
      axis.ticks.y=element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()) 
pdf("./Figures/PCoA/Ind319Hist.pdf",width=10,height=6)
BR
dev.off()
```

# Export PCoA Axes

```{r export_axes}
PCoAAxes<- PlotPcoa %>% dplyr::select(Pop,Lat, Num, V1, V2, V3, V4) %>% 
  rename(PCoA1 = V1,
         PCoA2 = V2,
         PCoA3 = V3,
         PCoA4 = V4) 
write.csv(PCoAAxes,"./Data/PCoAAxes.csv", row.names=F) 
```
